### 2015-07-16 2015-Ladutenko-Qabs/data/report-with-figures.tex
layered plot

%%\documentclass[fullscreen=true, handout]{beamer}
\documentclass[fullscreen=true]{beamer}
\usetheme[secheader]{Boadilla}
%%\usetheme{Boadilla}
%%\usetheme[height=7mm]{Rochester} 
%%\useinnertheme{rounded}
 \usecolortheme{seahorse}
%% \usecolortheme{whale}
%% \usecolortheme{rose}
%%\setbeamertemplate{navigation symbols}{\insertframenavigationsymbol}
%% or
\beamertemplatenavigationsymbolsempty
%% \setbeamertemplate[info line]{footline} %%Не работает
%% {\quad\strut\insertsection
%% \hfill\insertframenumber/\inserttotalframenumber\strut\quad} 
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} %%apt-get install cm-super
%%\usepackage[english, russian]{babel}
\usepackage[english]{babel}
\emergencystretch=35pt %%Чтобы нe было строчек, вылезших на поля
\usepackage{ragged2e} %% to use \justifying
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{listings}
%% \lstset{language=C++,
%%   backgroundcolor=\color{Black},
%%    basicstyle=\color{White}\tiny\ttfamily,
%%    keywordstyle=\color{Orange},
%%    identifierstyle=\color{Cyan},
%%    stringstyle=\color{Red}, 
%%    commentstyle=\color{Green},  
%%    breaklines=true,
%%    breakatwhitespace=true,
%%    tabsize=10,
%%    showstringspaces=false%%
%% }
%%or better
\lstset{breakatwhitespace,
language=C++,
columns=fullflexible,
keepspaces,
breaklines,
tabsize=2,
backgroundcolor=\color{white},
basicstyle=\color{black}\small\ttfamily,
keywordstyle=\color{orange},
identifierstyle=\color{blue},
stringstyle=\color{red}, 
commentstyle=\color{green},  
showstringspaces=false,
extendedchars=true}
\newcommand{\transhi}{40} %% Переменная - хорошая прозрачность
\setbeamercovered{transparent = \transhi} 
\setbeamertemplate{caption}[numbered]
\begin{document}
%% ----------------------------------------------------------------
\title{Report \today}
%% ----------------------------------------------------------------
\section{Figures}
\begin{frame}{Legend} %%Добавить интриги
  \begin{block}{}
    \begin{enumerate}
    \item blue - dipole
    \item red  - qudrupole
    \item green - octupole      
    \end{enumerate}
  \end{block}
  \begin{block}{}
    \begin{enumerate}
    \item solid - electric
    \item dashed  - magnetic
    \end{enumerate}
  \end{block}
  \begin{block}{}
    \begin{enumerate}
    \item thin - inner (core) layer
    \item thicker - middle layer
    \item thick - outher layer      
    \end{enumerate}
  \end{block}
All curves are normalized to it`s maxima! As maxima goes out ot the
plotted range normalization changes!!!
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+00.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+00.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+00.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+01.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+01.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+01.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+02.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+02.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+02.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+03.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+03.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+03.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+04.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+04.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+04.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+05.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+05.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+05.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+06.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+06.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+06.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+07.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+07.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+07.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+08.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+08.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+08.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+09.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+09.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+09.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+10.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+10.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+10.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+11.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+11.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+11.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+12.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+12.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+12.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+13.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+13.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+13.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+14.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+14.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+14.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+15.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+15.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+15.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+16.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+16.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+16.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+17.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+17.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+17.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+18.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+18.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+18.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+19.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+19.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+19.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+20.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+20.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+20.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+21.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+21.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+21.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+22.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+22.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+22.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+23.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+23.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+23.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+24.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+24.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+24.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+25.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+25.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+25.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+26.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+26.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+26.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+27.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+27.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+27.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+28.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+28.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+28.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+29.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+29.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+29.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+30.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+30.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+30.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+31.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+31.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+31.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+32.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+32.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+32.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+33.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+33.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+33.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+34.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+34.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+34.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+35.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+35.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+35.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+36.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+36.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+36.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+37.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+37.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+37.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+38.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+38.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+38.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+39.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+39.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+39.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+40.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+40.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+40.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+41.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+41.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+41.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+42.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+42.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+42.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+43.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+43.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+43.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+44.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+44.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+44.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+45.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+45.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+45.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+46.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+46.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+46.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+47.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+47.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+47.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+48.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+48.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+48.pdf}
  \end{figure}
\end{frame}

\subsection{2015-04-01-SiAgSi-ab-spectra-calc-d+49.pdf}
\begin{frame}
  \begin{figure}
    \includegraphics[height=0.9\textheight]{fig/2015-04-01-SiAgSi-ab-spectra-calc-d+49.pdf}%
    \caption{2015-04-01-SiAgSi-ab-spectra-calc-d+49.pdf}
  \end{figure}
\end{frame}
\end{document}