### 2015-07-22 2015-Ladutenko-Qabs/2015-Ladutenko-Qabs-ilya-v2.tex
Clean up

%\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress]{revtex4-1}  % for review and submission
%\documentclass[aps,preprint,showpacs,superscriptaddress,groupedaddress]{revtex4-1}  % for double-spaced preprint

%Phys Rev B
%Letter or rapid communication - 3500 words
%\documentclass[prb,groupedaddress,reprint]{revtex4-1}
%APL
\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress]{revtex4-1}
%\documentclass[aps,prl,reprint]{revtex4-1}
%\documentclass[aip,jap,reprint]{revtex4-1}
\usepackage[english]{babel}
\usepackage{graphicx}% Include figure files
\usepackage[pdftex,unicode,colorlinks, citecolor=blue,%
filecolor=black, linkcolor=blue, urlcolor=black]{hyperref}
\usepackage[figure,table]{hypcap} %links should lead to the begining
                                %of figure or table...

\begin{document} %Max size in pdf for APL - 4 print pages

% \title{More efficient than a ``super'' absorption in spherical
% nanoparticles} %KSL
% \title{Super-absorption and efficient absorption of light by
% spherical nanoparticles} %IVS
\title{Super absorption of light by nanoparticles} %AEM
% Proposed reviewers: Anatoly Zayats
% Andrea Alu
% Sergei Bozhevolnyi

\author{Konstantin Ladutenko} \email[e-mail: ]{fisik2000@mail.ru}
\affiliation{ITMO University, 49 Kronverskii Ave., St.~Petersburg
  197101, Russian Federation}

\affiliation{Ioffe Physical-Technical Institute of the Russian Academy
  of Sciences, 26 Polytekhnicheskaya Str., St.~Petersburg 194021,
  Russian Federation}

\author{Pavel Belov} \affiliation{ITMO University, 49 Kronverskii
  Ave., St.~Petersburg 197101, Russian Federation}

\author{Ovidio Pe\~{n}a-Rodr\'{i}guez} \affiliation{Instituto de
  Fusi\'{o}n Nuclear, Universidad Polit\'{e}cnica de Madrid, Jos\'{e}
  Guti\'{e}rrez Abascal 2, E-28006 Madrid, Spain}

\author{Ali Mirzaei} \author{Andrey Miroshnichenko} \author{Ilya
  Shadrivov} \affiliation{Nonlinear Physics Centre, Research School of
  Physics and Engineering, The Australian National University, 59
  Mills Rd, Acton, ACT, 2601, Australia}

\date{\today}
% APL 250 words! Phys Rev B less then 500 words, about
% 5% ot total paper length
% emacs M-x count-words

\begin{abstract}
  Nanoparticles have a fundamental limit as to how much light they can
  absorb. This limit is based on the finite number of modes excited
  inside the nanoparticle at a given wavelength and maximal absorption
  capacity per mode. The enhanced absorption can be achieved
  when each mode supported by the nanopartice absorbs light up to the
  maximum capacity. Using stochastic optimization algorithm, we design
  multi-layer nanoparticles, in which we can make several resonant
  modes overlap at the same frequency resulting in {\it
    super-absorption}.  We further introduce the {\it efficiency of
    the absorption} for a nanoparticle, which is the absorption normalized
  by the physical size of the particle, and show that efficient
  absorbers are not always operating in super-absorbing regime.
\end{abstract}

\pacs% insert suggested PACS numbers in braces on next line
{41.20.Jb 42.25.Bs 02.60.Pn 02.70.-c}
% 41.20.Jb Electromagnetic wave propagation 42.25.Bs Wave propagation,
% transmission and absorption
%% 42.25.Fx Diffraction and scattering
% 02.60.Pn Numerical optimization 02.70.-c Computational techniques;
% simulations

\maketitle %\maketitle must follow title, authors, abstract and \pacs

Mie theory~\cite{Mie-1908}, which is over 100 years old, describes
interaction of electromagnetic waves with spherical particles. Mie
solution is still of great interest these days~\cite{Suzuki-2008,MacKowski-2012,Lerme-2000,Xu-2005,Li-2006,Gogoi-2010,Santiago-2011},
since it is one of the primary tools for analyzing wave scattering by
spherical objects. Further development of the Mie
theory~\cite{Yang-2003, Pena-scattnlay-2009} made it possible to apply
it to study the multi-layered spherical
particles~\cite{Sheehan-2013,Selmke-2012}.  Such particles have
various applications in cancer treatment~\cite{Zhang-2010,
  Hirsch-2003}, medical diagnostics~\cite{Allain-2002},
cloaking~\cite{Qui-2009, Semouchkina-2013, Ladutenko-2014} and
plasmonic devices~\cite{Martin-2013, Alu-2005}, in the study of thermal properties of insulating materials~\cite{Xie-2013}, as well as for improving solar cells performance~\cite{Kameya-2011,Mann-2011}.

The scattering properties of multilayer cylinders and spheres was
studied in great detail by Ruan and Fan~\cite{Fan-2010,Fan-2011}.  In
these works authors introduced the concept of a super-scattering, when
the scattering cross-section of a multilayer particle exceeds that of
a homogeneous particle of the same size in the so-called
single-channel limit. The super-scattering appears when a
multi-layer structure has several nearly degenerate modes,
i.e. their resonance frequencies coincide or close to each other. In a homogeneous particle, the resonances appear at different frequencies, and there is no design freedom to make these resonances overlap, and this limits the
achievable scattering cross-section.

Similar fundamental limitations exist for the absorbing properties of
sub-wavelength nanoparticles.  Tribelsky~\cite{Tribelsky-2011} has
derived a theoretical limit of a maximum absorption cross-section
(ACS) value for a single channel, i.e., when only one mode of the
sphere is excited.  As a result the absorption coefficients
$\tilde{a}_n= {\rm Re}\{a_n\} - |a_n|^2 $ and $\tilde{b}_n= {\rm
  Re}\{b_n\} - |b_n|^2 $ become limited by $1/4$, here $a_n$ and $b_n$
are scattering coefficient as defined within the Mie
theory~\cite{Bohren-1983}.

To overcome these limitations, we employ similar approach to the one used for
enhancing absorption cross-section~\cite{Fan-2011}. In particular, we
propose to use the multi-layer structures, and by means of stochastic
optimization algorithm we optimize the ACS of such nanoparticles. We
analyze the absorption cross-section of these particles, and present
the super-absorption regime. We further introduce the absorption
efficiency, which is the ACS normalized to the geometric cross-section
of the particles. Here we show that there is a strong counter-play
between the increased absorption for larger particles vs size for
smaller particles, and quite remarkably we find that the most {\em
  efficient} absorption can be achieved in a single channel limit for
small particles.

Another approach for an ideal absorber is given in a recent work by
Grigoriev et al.~\cite{Grigoriev-2015}. Authors considered only
a dipole approximation, and the final result is very close to the dipole limit predicted by
Tribelsky~\cite{Tribelsky-2011}.  Such absorption design
corresponds to the single mode limit.  At the same time, Grigoriev et
al.~\cite{Grigoriev-2015} also provide an equation to design a
core-shell structure from given materials. However, in the case of
$Si$ core and $Ag$ shell materials and sizes taken from the best design obtained in the present paper, their equation gives a complex value for the core material filling factor, which cannot be achieved in an experimental realisations.

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/Fig1}%
    \caption{ Shematic view of the simulated $Si/Ag/Si$ particle. 
      \label{fig:geom}
    }%
  }
\end{figure}

We start our analyses by considering the triple layered $Si/Ag/Si$
spherical particle illuminated with a plane
wave schematically shown in Fig.~\ref{fig:geom}. In what follows we describe the materials
using experimentally measured parameters from the
Ref.~\cite{palik-1997}, and, e.g. at $\lambda = 500$ nm $\epsilon_{Si} = 18.5 + i0.63$ and
$\epsilon_{Ag} = -8.5 + i0.76$.  To optimize the thickness of each
layer we implemented~\cite{JADE-web} an adaptive differential
evolution algorithm~\cite{Storn-DE-first-1997}, which is called
JADE~\cite{Jingqiao-JADE-2009}.  The technical details of the
optimization algorithm are published previously in
Ref.~\cite{Ladutenko-2014}. We perform Mie calculations using the
Scattnlay software~\cite{Pena-scattnlay-2009,Scattnlay-web}, whose
results are verified by a number of other implementations of the
Mie solutions and by a commercially available software including CST
Microwave studio~\cite{CST-web} and Comsol Multiphysics~\cite{Comsol-web}.

It is a common understanding that, in general, a bigger particle 
has a bigger absorption cross-section, so sphere with the diameter of
1 cm absorbs more light than any nanoscale sphere. Therefore, it is practical to employ {\em the absorption
efficiency} $Q_{\rm abs}=C_{\rm abs}/\pi R^2$, where $R$ is the outer
radius of the particle and $C_{\rm abs}$ is the absorption
cross-section.

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/2015-04-01-Qabs-SiAgSi-overview}%
    \caption{ Results of the optimization of the absorption efficiency
      for the fixed wavelength of 500~nm. (a) Absorption efficiency
      with the best value achieved for the particle of the radius of 36~nm and
      Ag/Si design (zero sized core). Dashed lines show theoretical
      limits for the first channel and second channel
      absorption. Second and third peaks in the absorption efficiency
      curve exceed the theoretical limit for the second mode
      absorption at $R=63$~nm and $R=81$~nm. Local maxima of
      absorption efficiency are additionally marked with arrows. (b) Mie absorption
      coefficients for individual excited modes of the optimized
      structures. (c) Optimized layer thicknesses. For the total
      particle radius below 46~nm the optimizer converges to
      the two-layer structure, when core size vanishes, and the
      optimum design is a bi-layer $Ag/Si$ particle. 
      \label{fig:overview}
    }%
  }
\end{figure}
%
In order to study the dependence of the absorption efficiency on the
outer particle size, we run optimization algorithm for different
(fixed) particle outer size, and our optimization parameters are the
radii of internal cores, whereas the target function is the absorption
efficiency. We maximize absorption efficiency at a fixed wavelength
of the incident light (we have chosen $\lambda=500$~nm). We show the results of our stochastic optimization
algorithm in Fig.~\ref{fig:overview}~(a).  Dashed lines show
theoretical absorption limit of a dipole ($n=1$) and a quadrupole
($n=2$) resonances~\cite{Tribelsky-2011}, which are given
as $$Q^{(n)}_{\rm abs\ max}=\frac{2n+1}{2q^2},$$ where the size
parameter $q=2\pi R/\lambda$, and $n$ is the angular momentum of the
mode. Following Ref.~\cite{Fan-2011}, where authors introduce
super-scattering for spherical particles, here we we introduce
super-absorption regime, when {\em the ACS is larger than the theoretical limit
for absorption by the mode with highest excited angular momentum
$n$}. In our parameter space we have just modes up to the quadrupole
excited ($n=2$), and in order to get a super-absorption our efficiency
should be higher than that of a quadrupole. We clearly see this
super-absorption regime at $R>60$~nm, in Fig.~\ref{fig:overview}~(a).

In Fig.~\ref{fig:overview}~(b) we present the values of Mie absorption
coefficients for individual excited modes in the structure, while
horizontal dashed line shows the theoretical limit (1/4) for each of
them. $\tilde{a}_{1,2}$ are electric dipole and electric quadrupole,
while $\tilde{b}_{1,2}$ are magnetic dipole and magnetic
quadrupole. For small particles, as expected, the absorption is
dominated by an electric dipole $\tilde{a}_1$.  At $R > 56.6$~nm the
optimization procedure finds that the designs with both electric and
magnetic dipoles have larger ACS, than the structure with only the
electric dipole excited. This is why the curves in
Figs.~\ref{fig:overview}~(b,c) experience the discontinuity. We also
note that there is a very narrow range of particle sizes, between
80.7~nm and 82.1~nm, where our analyses finds that the design
supporting electric dipole $\tilde{a}_1$ and magnetic quadrupole
$\tilde{b}_2$ has larger ACS, and this explains two more
discontinuities of the curves at the respective size values.

Fig.~\ref{fig:overview}~(c) shows optimized sizes of the layers inside
the multi-layer structure. It reveals quite a curious result, that the
dipole branch (i.e. for particle radii below 56.6~nm) has two
parts. For $R<46$~nm the best absorber has just two layers, as the
radius of the core of the three-layer structure vanishes, and the
particle reduces to $Ag/Si$ core-shell structure.  At $R=46$~nm dipole
channel reaches its theoretical limit (it becomes
$\tilde{a}_1>0.249$).  It appears that the optimizer introduced the
inner $Si$ layer in order to keep $\tilde{a}_1$ near the theoretical
limit as the $R$ increases.  As a side effect, the quadrupole
contribution $\tilde{a}_2$ appears; however, it does not help to reach
super-absorption limit for $n=2$.

Remarkably, the absolute maximum {\em absorption efficiency} is not reached
within the super-absorption regime. Figure~\ref{fig:overview} shows
that the maximum efficiency is achieved for small particle size, and
the ACS is still well below the single channel limit. It appears that
the $Ag/Si$ core-shell nanoparticle, with the total radii of
approximately 36~nm is the most efficient absorber in the considered
structure, whose ACS reaches values {\em over 5 times the physical
cross-section area of the particle}.  From practical point of view it
is quite important that the maximum can be reached in a bi-layer
structure, instead of a triple-layer, since it should be easier and
cheaper to fabricate.

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/2015-04-01-SiAgSi-ab-spectra3.pdf}%
    \caption{Spectra of Mie absorption coefficients of (a) efficient and (b-c)super-absorption design.      
      \label{fig:spectra}
    }%
  }
\end{figure}
To study spectral properties of the structures with large ACS which we
obtained by the optimization, in Fig.~\ref{fig:spectra} we plot three
different cases for designs that correspond to local maxima of $Q_{\rm
  abs}$ shown in Fig.~\ref{fig:overview}~(a).  As expected, the design
corresponding to the maximum absorption efficiency at $R=36$~nm has a
single electric dipole resonance centered at the target wavelength
$\lambda=500$~nm. Spectra of designs with maxima at $R=63$~nm and
$R=81$~nm have a signature of the super-absorption, i.e. there is an
overlap of several resonances.  We note that these structures have
additional absorption resonances, but they are located far from the
wavelength of interest.

A noticeable feature of Fig.~\ref{fig:spectra}~(c) is an almost flat
top of the electric dipole resonance.  
%This effect might be related with the coupling between the electric
%dipole $\tilde{a}_1$ with the magnetic quadrupole $\tilde{b}_2$, but
%this hypothesis should be further tested.  Whereas resonant responses
%of other designs originate from coupling of incident plane wave with
%the corresponding multipole, it is not the case for the design with
%$R=81$~nm. Here, the particle is mostly composed from the $Si$ outer
%shell, that give enough volume for $\tilde{b}_2$.  Dipole response
%comes from the small inner part of the sphere and can not be directly
%excited with the incident wave.  It takes the power from the
%surrounding $\tilde{b}_2$ mode and very soon it reaches the
%fundamental limit for absorption.  This way, we can observe a flat top
%for the $\tilde{a}_1$ response accompanied with a significant value of
%$\tilde{b}_2$.  Our suggestion is indirectly proven by the reduced
%width of the $\tilde{b}_2$ resonance compared to other quadrupole
%responses.

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.45\textwidth]{fig/SiAgSi-flow-R62-YZ-Eabs.pdf}%
    \caption{ Amplitude of electric field for $R=36$~nm (a,c) and
      super-absorbing designs (b,d) in E-k  (a-b) and H-k (c-d)
      planes normalized to incident wave. 
      \label{fig:field}
    }%
  }
\end{figure}
Finally, we present distribution of the amplitude of the electric
field in Fig.~\ref{fig:field} for two designs: with the best
efficiency at $R=36$~nm and in a super-absorbing design with
$R=63$~nm.  Using semi-transparent white curves we also plot
streamlines of the Poynting vector which characterize energy flow.
For the effective design of the absorber, the power from a large
cross-sectional area flows into the particle.  In case of
super-absorption regime, we observe the formation of vortices in the
power flow, which make absorption more efficient as the
electromagnetic energy propagates several times through the absorbing
materials.  The reason for the smaller overall absorption efficiency
of a super-absorbing design is obvious: spatial distribution of the electric
field is not uniform inside the particle and the share of the volume
with high absorption rate in the vortices do not compensate low
absorption efficiency in the regions with small electric field (
absorbed power is expressed as $P_{\rm abs}=\frac{1}{2}\int\sigma
\left|E\right|^2dx\,dy\,dz$)

In conclusion, we introduce and study the effect of the super-absorption, when the
absorption cross-section of the nanoparticle can reach the theoretical
limit for several modes at the same frequency. This becomes possible when several resonant modes of the
structure overlap at the same frequency, and this regime can be achieved in multi-layer nanoparticles. Moreover, quite unexpectedly we find that the most efficient absorbers, which are characterized by
enhanced absorption efficiencies, are smaller nanoparticles working in
a single mode regime. We present their spectral characteristics and
field structure.  The stochastic algorithm utilized in our approach is
very generic and the optimization can be repeated for any desired
wavelength or wavelength range.  As a result, one can design
spectrally-selective absorbers or broadband absorbers with almost
arbitrary prescribed properties.  It is interesting to note
that a similar conclusion was made by Miller et al.~\cite{Miller-2014}
for extinction of arbitrary particles: small size with only dipole
response is preferable for geometric volume normalized efficiency.

\begin{acknowledgments}

 AEM and IVS would like to acknowledge support from the Australian Research Council through Future Fellow and Discovery project schemes. 
  % The authors would like to thank the Ministry of Education and
  % Science of the Russian Federation (Goszadanie 2014/190, Zadanie
  % No. 3.561.2014/K, project 14.584.21.0009 10), Russian Foundation
  % for Basic Research (Grant 15-02-01344), and Government of the
  % Russian Federation (Grant 074-U01) for the financial support. The
  % simulations of cloak designs has been funded by the Russian
  % Science Foundation Grant No. 14-12-01227.
\end{acknowledgments}

\bibliography{2015-Ladutenko-Qabs}
\end{document}